On the uniqueness and stability of dictionaries \\ for sparse representation of noisy signals

### 2019-05-10 manuscripts/IEEE/final_submission.tex
chaz

Before moving on to our proof of Thm.~\ref{SLCopt}, we briefly revisit here our discussion on sample complexity from the end of the previous section. While an exponential number of samples may very well prove to be necessary in the deterministic or almost-certain case, our proof of Thm.~\ref{DeterministicUniquenessTheorem} can be extended to hold with some probability for \emph{any} number of samples by alternative appeal to a probabilistic pigeonholing at the point early in the proof where the (deterministic) pigeonhole principle is applied to show that for every $S \in \mathcal{H}$, there exist $k$ vectors $\mathbf{x}_i$ supported on $S$ whose corresponding $\mathbf{\overline x}_i$ all share the same support.\footnote{A famous example of such an argument is the counter-intuitive ``birthday paradox", which demonstrates that the probability of two people having the same birthday in a room of twenty-three is greater than 50\%.}

Given insufficient samples, this argument has some less-than-certain probability of being valid for each $S \in \mathcal{H}$. Nonetheless, simulations with small hypergraphs confirm that the probability of success very quickly approaches one once the number of samples $N$ surpasses a small fraction of the deterministic sample complexity.

@@

We close with some challenges left open by our work. All conditions stated here guaranteeing the uniqueness and stability of sparse representations have only been shown sufficient; it remains open, therefore, to extend them to necessary conditions, be they on required sample size, the structure of support set hypergraphs, or tolerable error bounds. On this last note, we remark that our deterministic treatment considers always the ``worst-case" noise, whereas the ``effective" noise sampled from a concentrated distribution might be significantly reduced, especially for high-dimensional data. It would be of great practical benefit to see how drastically all conditions can be relaxed by requiring only probabilistic guarantees in this way, or in the spirit of the discussion on probabilistic pigeonholing for reducing sample complexity (as in the famous ``birthday paradox") following our proof of Thm.~\ref{DeterministicUniquenessTheorem} .

Another interesting question raised by our work is for which special cases is it efficient to check that a solution to Prob.~\ref{InverseProblem} or \ref{OptimizationProblem} is unique up to noise and inherent ambiguities. Considering that the sufficient conditions we have described for checking this in general are NP-hard to compute, are the necessary conditions hard? Are Probs.~\ref{InverseProblem} and \ref{OptimizationProblem} then also hard \cite{Tillmann15}? Finally, since solving Prob.~\ref{SLCopt} directly is intractable in general, but efficiently solvable by $\ell_1$-norm minimization for fixed matrices with large enough lower bounds over sparse domains \cite{eldar2012compressed}, is there a version of Thm.~\ref{SLCopt} certifying when Prob.~\ref{OptimizationProblem} can be solved efficiently in a similar way?  % [*** Incorporate this open problem with more words ***]

@@

\textbf{Acknowledgement.} We thank F. Sommer and D. Rhea for early thoughts, and I. Morris for posting \eqref{eqdim} online.

>>>>

Before moving on to our proof of Thm.~\ref{SLCopt}, we briefly revisit here our discussion on sample complexity from the end of the previous section. While an exponential number of samples may very well prove to be necessary in the deterministic or almost-certain case, our proof of Thm.~\ref{DeterministicUniquenessTheorem} can be extended to hold with some probability for \emph{any} number of samples by alternative appeal to a probabilistic pigeonholing at the point early in the proof where the (deterministic) pigeonhole principle is applied to show that for every $S \in \mathcal{H}$, there exist $k$ vectors $\mathbf{x}_i$ supported on $S$ whose corresponding $\mathbf{\overline x}_i$ all share the same support.\footnote{A famous example of such an argument is the counter-intuitive ``birthday paradox", which demonstrates that the probability of two people having the same birthday in a room of twenty-three is greater than 50\%.} 
Given insufficient samples, this argument has some less-than-certain probability of being valid for each $S \in \mathcal{H}$. Nonetheless, simulations with small hypergraphs confirm that the probability of success very quickly approaches one once the number of samples $N$ surpasses a small fraction of the deterministic sample complexity.

@@

We close with some challenges left open by our work. All conditions stated here guaranteeing the uniqueness and stability of sparse representations have only been shown sufficient; it remains open, therefore, to extend them to necessary conditions, be they on required sample size, the structure of support set hypergraphs, or tolerable error bounds. On this last note, we remark that our deterministic treatment considers always the ``worst-case" noise, whereas the ``effective" noise sampled from a concentrated distribution might be significantly reduced, especially for high-dimensional data. It would be of great practical benefit to see how drastically all conditions can be relaxed by requiring only probabilistic guarantees in this way, or in the spirit of our discussion on probabilistic pigeonholing to reduce sample complexity (as in the famous ``birthday paradox") following the proof of Thm.~\ref{DeterministicUniquenessTheorem}.

Another interesting question raised by our work is for which special cases is it efficient to check that a solution to Prob.~\ref{InverseProblem} or \ref{OptimizationProblem} is unique up to noise and inherent ambiguities. Considering that the sufficient conditions we have described for checking this in general are NP-hard to compute, are the necessary conditions hard? Are Probs.~\ref{InverseProblem} and \ref{OptimizationProblem} then also hard (e.g., see \cite{Tillmann15})? Finally, since Prob.~\ref{SLCopt} is intractable in general, but efficiently solvable by $\ell_1$-norm minimization when the matrix is known (and has a large enough lower bound over sparse domains \cite{eldar2012compressed}), is there a version of Thm.~\ref{SLCopt} certifying when Prob.~\ref{OptimizationProblem} can be solved efficiently in full by similar means?  % [*** Incorporate this open problem with more words ***]

@@

\textbf{Acknowledgement.} We thank Fritz Sommer and Darren Rhea for early thoughts, and Ian Morris for posting \eqref{eqdim} online.
